\section{OpenBDLM main menu}
\label{S:OPENBDLMMAINMENU}

Once a project is loaded, the OpenBDLM main menu (see Listing~\ref{LST:OpenBDLMMainMenu}) displays the list of actions which can be done. 
The OpenBDLM main menu appears each time a selected action is done; until the user types \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!Q!} to save the project and quit the program.

\begin{itemize}
    \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!1!}: estimates the model parameters
    \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!2!}: estimates the initial hidden states
    \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!3!}: estimate the hidden states
    \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!11!}: display the current model parameters values on the \MATLAB{} command line window and allows the user to modify the model parameters values
     \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!12!}: display the initial hidden states values on the \MATLAB{} command line window and allows the user to modify the initial hidden states values
     \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!13!}: display information about the current training period  on the \MATLAB{} command line window and allows the user to modify the training period
     \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!14!}: plot figures on the screen about the data in memory, the data availability, and the hidden states estimation results
     \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!15!}: display the content of the model matrices on the \MATLAB{} command line window at a specific time
     \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!16!}: create synthetic data from the current information in memory
     \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!17!}: export data and results as well as the figures in PNG, PDF and  \LaTeX{}   format
     \item option  \colorbox{light-gray}{\lstinline[basicstyle = \mlttfamily \small ]!18!}: display the current options 
    \end{itemize}

\begin{lstlisting}[ frame = single, basicstyle = \mlttfamily \small, caption = { OpenBLDM main menu}, label = LST:OpenBDLMMainMenu,  float =h!, linewidth=\linewidth, captionpos=b]
---------------------------------------------
/    Choose from 
--------------------------------------------- 

     1  ->  Learn model parameters values 
     2  ->  Estimate initial hidden states values 
     3  ->  Estimate hidden states values 

     11 ->  Display and modify current model parameter values 
     12 ->  Display and modify current initial hidden states values 
     13 ->  Display and modify current training period 
     14 ->  Plots 
     15 ->  Display model matrices 
     16 ->  Create synthetic data 
     17 ->  Export
     18 ->  Display current options in configuration file format 

     Type Q to Save and Quit 

     choice >> 
\end{lstlisting}